Signatures and Induction Principles for Higher Inductive-Inductive Types

### 2019-01-30 paper.tex
Morphisms progress

\newcommand{\Elim}{\mathsf{Elim}}
\newcommand{\elim}{\mathsf{elim}}
\newcommand{\Rec}{\mathsf{Rec}}
\newcommand{\record}{\mathsf{record}}
\newcommand{\funext}{\mathsf{funext}} \newcommand{\Q}{\mathsf{Q}}
\newcommand{\T}{\mathsf{T}} \newcommand{\leaf}{\mathsf{leaf}}
\newcommand{\node}{\mathsf{node}} \newcommand{\perm}{\mathsf{perm}}
\newcommand{\coe}{\mathsf{coe}} \newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}} \newcommand{\untr}{\mathsf{untr}}
\newcommand{\from}{\mathsf{from}}
\newcommand{\fromeq}{{\mathsf{from}\hspace{-0.3em}\equiv}}
\newcommand{\fromsimeq}{{\mathsf{from}\hspace{-0.3em}\simeq}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\DModel}{\mathsf{DModel}}
\newcommand{\module}{\mathsf{module}}
\newcommand{\open}{\mathsf{open}}
\renewcommand{\P}{\mathsf{P}} \newcommand{\Bool}{\mathsf{Bool}}
\newcommand{\true}{\mathsf{true}} \newcommand{\false}{\mathsf{false}}
\renewcommand{\not}{\mathsf{not}} \newcommand{\0}{\mathsf{0}}
\newcommand{\1}{\mathsf{1}} \renewcommand{\Pr}{\mathsf{Pr}}
\newcommand{\PrNat}{\mathsf{PrNat}} \newcommand{\J}{\mathsf{J}}
\newcommand{\wkV}{\mathsf{wkV}} \renewcommand{\r}[1]{{\P_{#1}}}
\newcommand{\stab}{\mathsf{stab}} \newcommand{\NTy}{\mathsf{NTy}}
\newcommand{\isDec}{\mathsf{isDec}} \newcommand{\dec}{\mathsf{dec}}
\newcommand{\yes}{\mathsf{yes}} \newcommand{\no}{\mathsf{no}}
\newcommand{\case}{\mathsf{case}} \newcommand{\inj}{\mathsf{inj}}
\newcommand{\K}{\mathsf{K}} \newcommand{\lb}{\langle}
\newcommand{\rb}{\rangle} \newcommand{\Decl}{\mathsf{Decl}}
\newcommand{\Core}{\mathsf{Core}} \newcommand{\IND}{\mathsf{ind}}
\newcommand{\Id}{\mathsf{Id}} \newcommand{\Base}{\mathsf{Base}}
\newcommand{\transport}{\mathsf{transport}}
\newcommand{\I}{\mathsf{I}} \newcommand{\E}{\mathsf{S}}
\newcommand{\transp}{\mathsf{transp}}
\newcommand{\Transp}{\mathsf{Transp}}
\newcommand{\W}{\mathsf{W}}
\newcommand{\Fin}{\mathsf{Fin}} \newcommand{\fzero}{\mathsf{fzero}}
\newcommand{\fsuc}{\mathsf{fsuc}}
\newcommand{\inv}{\mathsf{inv}}
\newcommand{\con}{\mathsf{con}}
\newcommand{\LEFT}{\mathsf{left}}
\newcommand{\RIGHT}{\mathsf{right}}
\newcommand{\seg}{\mathsf{seg}}
\newcommand{\Int}{\mathsf{Int}}
\renewcommand{\in}{\mathbin{\hat:}}
\renewcommand{\hat}[1]{{\color{BrickRed}{#1}}}
\newcommand{\blc}[1]{{\color{Black}{#1}}}
\newcommand{\vdashh}{\mathbin{\hat\vdash}}
\newcommand{\rah}{\mathbin{\hat\ra}}
\newcommand{\commah}{\hat,\,}
\newcommand{\timesh}{\mathbin{\hat\times}}
\newcommand{\eqh}{\mathbin{\hat=}}
\newcommand{\TR}{\hat{\mathsf{tr}}}
\newcommand{\ap}{\hat{\mathsf{ap}}}
\newcommand{\apd}{\hat{\mathsf{apd}}}
\renewcommand{\tt}{\hat{\mathsf{tt}}}
\newcommand{\Tel}{\hat{\mathsf{Tel}}}
\newcommand{\Type}{\hat{\mathsf{Type}}}
\newcommand{\emptytel}{\hat{\epsilon}}
\newcommand{\telext}{\mathbin{\hat{\lhd}}}
\newcommand{\ltel}{\hat{\ll}}
\newcommand{\rtel}{\hat{\rr}}
\newcommand\pp{\ensuremath{\hat{\mathbin{+\mkern-10mu+}}}}
\newcommand{\semicol}{\hat;\,}
\newcommand{\targetass}{\hat{\Gamma}\semicol}
\newcommand{\Alg}{\mathsf{Alg}}
\newcommand{\DisplayedAlg}{\mathsf{DisplayedAlg}}
\newcommand{\Morphism}{\mathsf{Morphism}}
\newcommand{\Section}{\mathsf{Section}}
\newcommand{\ass}{\mathsf{ass}}
\newcommand{\idl}{\mathsf{idl}}
\newcommand{\idr}{\mathsf{idr}}

@@

We will use the following functions defined using $\hat{\J}$ in the
standard way. We write $\TR_{\hat{P}}\,\hat{e}\,\hat{t}\in
\hat{P}\,\hat{v}$ for transport of $\hat{t} \in \hat{P}\,\hat{u}$
along $\hat{e} \in \hat{u}\eqh \hat{v}$. We write
$\ap\,\hat{f}\,\hat{e}\in\hat{f}\,\hat{u}\eqh\hat{f}\,\hat{v}$ for
$\hat{f}:\hat{A}\rah\hat{B}$ and $\hat{e}\in\hat{u}\eqh\hat{v}$,
$\hat{\apd\,\hat{f}\,\hat{e} \in \TR_{\hat{P}}\,\hat{e}\,(\hat{f}\,\hat{u})
\eqh \hat{f}\,\hat{v}}$ for $\hat{\hat{f}\in(\hat{x}\in \hat{A})\rah
\hat{B}}$ and $\hat{e} \in \hat{u}\eqh \hat{v}$.

@@

Note that this singleton contraction is not possible for a general
non-strictly positive function space. We rely on the domain being small:
$(\underline{a})^\M$ yields an equation, but for general $\Gamma\vdash A$ types,
$A^\M$ only yields an unknown relation.

\subsection{The full homomorphism interpretation}

TDAODKAS

@@

\[
\infer{\hat{\Gamma}\vdashh t^\D \in \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\hat{(}\hat{\gamma^D}\in\Delta^\D\,\hat{\gamma}\hat{)}\rah  A^\D\,\hat{\gamma}\,\hat{\gamma^D}\,\hat{(}t^\A\,\hat{\gamma}\hat{)}}{\hat{\Gamma}\semicol\Delta\vdash t : A}
\]
We introduce the following shorthand: $t\,\hat{\gamma}\,\hat{\gamma^D}$ is
abbreviated as $t\,\hat{\gamma^2}$ for some $t$ expression. The
implementation of $\blank^\D$ is given below.  \begingroup
%% \allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\D\,\hat{\gamma} && :\equiv \hat{\top} \\
  & (\Delta,\,x:A)^\D\,(\hat{\gamma}\commah\,\hat{t}) && :\equiv \hat{(}\hat{\gamma^D}\in\Delta^\D\,\hat{\gamma}\hat{)}\timesh A^\D\,\hat{\gamma^2}\,\hat{t} \\
  & x^\D\,\hat{\gamma^2} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^D} \\ % TODO: De Bruijn
  & \U^\D\,\hat{\gamma^2}\,\hat{A} && :\equiv \hat{A} \rah  \Type_{\hat{i}} \\
  & (\underline{a})^\D\,\hat{\gamma^2}\,\hat{t} && :\equiv a^\D\,\hat{\gamma^2}\,\hat{t} \\
  & ((x:a)\ra B)^\D\,\hat{\gamma^2}\,\hat{f} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\hat{(}\hat{x^D}\in a^\D\,\hat{\gamma^2}\,\hat{x}\hat{)} \rah  B^\D\,\hat{(}\hat{\gamma}\commah\hat{x}\hat{)}\,\hat{(}\hat{\gamma^D}\commah\hat{x^D}\hat{)}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,u)^\D\,\hat{\gamma^2} && :\equiv \hat{(}t^\D\,\hat{\gamma^2}\hat{)}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\,\hat{(}u^\D\,\hat{\gamma^2}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\D\,\hat{\gamma^2}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  B^\D\,\hat{\gamma^2}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\D\,\hat{\gamma^2} && :\equiv t^\D\,\hat{\gamma^2}\,\hat{u} \\
  & (t=_a u)^\D\,\hat{\gamma^2}\,\hat{e} && :\equiv \TR_{\hat{(}a^\D\,\hat{\gamma^2}\hat{)}}\,\hat{e}\,\hat{(}t^\D\,\hat{\gamma^2}\hat{)} \eqh u^\D\hat{\gamma^2} \\
  & (\refl_t)^\D\,\hat{\gamma^2} && :\equiv \hat{\refl}_{\hat{(}t^\D\,\hat{\gamma^2}\hat{)}} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\D\,\hat{\gamma^2} && :\equiv \hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}pr^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)}\hat{\big)}\,\hat{(}eq^\D\,\hat{\gamma^2}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\D\,\hat{\gamma^2} && :\equiv \hat{\refl} \\
  & ((\hat{x}\in \hat{A})\ra b)^\D\,\hat{\gamma^2}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  b^\D\,\hat{\gamma^2}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\D\,\hat{\gamma^2} && :\equiv t^\D\,\hat{\gamma^2}\,\hat{u}
\end{alignat*}\endgroup
The predicate for a context is given by iterating $\blank^\D$ for its
constituent types. For a variable, the corresponding witness is looked up from
$\hat{\gamma^D}$.

@@

In addition to $\hat{\gamma^2}$, we use $t\,\hat{\gamma^3}$ to abbreviate $t\,\hat{\gamma}\,\hat{\gamma^D}\,\hat{\gamma^S}$. The implementation is the following.

@@

  & (\Delta,\,x:A)^\S\,(\hat{\gamma},\hat{t})\,(\hat{\gamma^D},\hat{t^D}) && :\equiv \hat{(}\hat{\gamma^S}\in\Delta^\S\hat{\gamma^2}\hat{)}\timesh A^\S\,\hat{\gamma^3}\,\hat{t}\,\hat{t^D} \\
  & x^\S\,\hat{\gamma^3} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^S} \\ % TODO: De Bruijn
  & \U^\S\,\hat{\gamma^3}\,\hat{A}\,\hat{A^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  \hat{A^D}\,\hat{x} \\
  & (\underline{a})^\S\,\hat{\gamma^3}\,\hat{t}\,\hat{t^D} && :\equiv a^\S\,\hat{\gamma^3}\,\hat{t} \eqh \hat{t^D} \\
  & ((x:a)\ra B)^\S\,\hat{\gamma^3}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\S\,\hat{(}\hat{\gamma,x}\hat{)}\,\hat{(}\hat{\gamma^D,}\,a^\S\,\hat{\gamma^3}\,\hat{x}\hat{)}\,\hat{(}\hat{\gamma^S,}\,\hat{\refl}\hat{)} \\
  & && \hspace{8.4em} \hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{\big(}\hat{f^D}\,\hat{x}\,\hat{(}a^\S\,\hat{\gamma^3}\,\hat{x}\hat{)}\hat{\big)} \\
  & (t\,u)^\S\,\hat{\gamma^3} && :\equiv \hat{\J}\,\hat{(}t^\S\,\hat{\gamma^3}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\hat{)}\,\hat{(}u^\S\,\hat{\gamma^3}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\S\,\hat{\gamma^3}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah B^\S\,\hat{\gamma^3}\,\hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{(}\hat{f^D}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^3} && :\equiv t^\S\,\hat{\gamma^3}\,\hat{u} \\
  & (t=_a u)^\S\,\hat{\gamma^3}\,\hat{e} && :\equiv \TR\,\hat{(}t^\S\,\hat{\gamma^3}\hat{)}\hat{\big(}\TR\,\hat{(}u^\S\,\hat{\gamma^3}\hat{)}\,\hat{(}\apd\,\hat{(}a^\S\,\hat{\gamma^3}\hat{)}\,\hat{e}\hat{)}\hat{\big)} \\
  & (\refl_t)^\S\,\hat{\gamma^3} && :\equiv \hat{\J}\,\hat{\refl}\,\hat{(}t^\S\,\hat{\gamma^3}\hat{)} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\S\,\hat{\gamma^3} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\bigg(}\hat{\J}\,\hat{\Big(}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\,\hat{pr^D}\,\hat{pr^S}.\,\hat{pr^S}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^3}\hat{)}\,\\

>>>>

\newcommand{\Elim}{\mathsf{Elim}}
\newcommand{\elim}{\mathsf{elim}}
\newcommand{\Rec}{\mathsf{Rec}}
\newcommand{\record}{\mathsf{record}}
\newcommand{\funext}{\mathsf{funext}} \newcommand{\Q}{\mathsf{Q}}
\newcommand{\T}{\mathsf{T}} \newcommand{\leaf}{\mathsf{leaf}}
\newcommand{\node}{\mathsf{node}} \newcommand{\perm}{\mathsf{perm}}
\newcommand{\coe}{\mathsf{coe}} \newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}} \newcommand{\untr}{\mathsf{untr}}
\newcommand{\from}{\mathsf{from}}
\newcommand{\fromeq}{{\mathsf{from}\hspace{-0.3em}\equiv}}
\newcommand{\fromsimeq}{{\mathsf{from}\hspace{-0.3em}\simeq}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\DModel}{\mathsf{DModel}}
\newcommand{\module}{\mathsf{module}}
\newcommand{\open}{\mathsf{open}}
\renewcommand{\P}{\mathsf{P}} \newcommand{\Bool}{\mathsf{Bool}}
\newcommand{\true}{\mathsf{true}} \newcommand{\false}{\mathsf{false}}
\renewcommand{\not}{\mathsf{not}} \newcommand{\0}{\mathsf{0}}
\newcommand{\1}{\mathsf{1}} \renewcommand{\Pr}{\mathsf{Pr}}
\newcommand{\PrNat}{\mathsf{PrNat}} \newcommand{\J}{\mathsf{J}}
\newcommand{\wkV}{\mathsf{wkV}} \renewcommand{\r}[1]{{\P_{#1}}}
\newcommand{\stab}{\mathsf{stab}} \newcommand{\NTy}{\mathsf{NTy}}
\newcommand{\isDec}{\mathsf{isDec}} \newcommand{\dec}{\mathsf{dec}}
\newcommand{\yes}{\mathsf{yes}} \newcommand{\no}{\mathsf{no}}
\newcommand{\case}{\mathsf{case}} \newcommand{\inj}{\mathsf{inj}}
\newcommand{\K}{\mathsf{K}} \newcommand{\lb}{\langle}
\newcommand{\rb}{\rangle} \newcommand{\Decl}{\mathsf{Decl}}
\newcommand{\Core}{\mathsf{Core}} \newcommand{\IND}{\mathsf{ind}}
\newcommand{\Id}{\mathsf{Id}} \newcommand{\Base}{\mathsf{Base}}
\newcommand{\transport}{\mathsf{transport}}
\newcommand{\I}{\mathsf{I}} \newcommand{\E}{\mathsf{S}}
\newcommand{\transp}{\mathsf{transp}}
\newcommand{\Transp}{\mathsf{Transp}}
\newcommand{\W}{\mathsf{W}}
\newcommand{\Fin}{\mathsf{Fin}} \newcommand{\fzero}{\mathsf{fzero}}
\newcommand{\fsuc}{\mathsf{fsuc}}
\newcommand{\inv}{\mathsf{inv}}
\newcommand{\con}{\mathsf{con}}
\newcommand{\LEFT}{\mathsf{left}}
\newcommand{\RIGHT}{\mathsf{right}}
\newcommand{\seg}{\mathsf{seg}}
\newcommand{\Int}{\mathsf{Int}}
\renewcommand{\in}{\mathbin{\hat:}}
\renewcommand{\hat}[1]{{\color{BrickRed}{#1}}}
\newcommand{\blc}[1]{{\color{Black}{#1}}}
\newcommand{\vdashh}{\mathbin{\hat\vdash}}
\newcommand{\rah}{\mathbin{\hat\ra}}
\newcommand{\commah}{\hat,\,}
\newcommand{\timesh}{\mathbin{\hat\times}}
\newcommand{\eqh}{\mathbin{\hat=}}
\newcommand{\TR}{\hat{\mathsf{tr}}}
\newcommand{\ap}{\hat{\mathsf{ap}}}
\newcommand{\apd}{\hat{\mathsf{apd}}}
\renewcommand{\tt}{\hat{\mathsf{tt}}}
\newcommand{\Tel}{\hat{\mathsf{Tel}}}
\newcommand{\Type}{\hat{\mathsf{Type}}}
\newcommand{\emptytel}{\hat{\epsilon}}
\newcommand{\telext}{\mathbin{\hat{\lhd}}}
\newcommand{\ltel}{\hat{\ll}}
\newcommand{\rtel}{\hat{\rr}}
\newcommand\pp{\ensuremath{\hat{\mathbin{+\mkern-10mu+}}}}
\newcommand{\semicol}{\hat;\,}
\newcommand{\targetass}{\hat{\Gamma}\semicol}
\newcommand{\Alg}{\mathsf{Alg}}
\newcommand{\DisplayedAlg}{\mathsf{DisplayedAlg}}
\newcommand{\Morphism}{\mathsf{Morphism}}
\newcommand{\Section}{\mathsf{Section}}
\newcommand{\ass}{\mathsf{ass}}
\newcommand{\idl}{\mathsf{idl}}
\newcommand{\idr}{\mathsf{idr}}
\newcommand{\inv}{\mathsf{inv}}

@@

We will use the following functions defined using $\hat{\J}$ in the standard
way. We write $\TR_{\hat{P}}\,\hat{e}\,\hat{t}\in \hat{P}\,\hat{v}$ for
transport of $\hat{t} \in \hat{P}\,\hat{u}$ along $\hat{e} \in \hat{u}\eqh
\hat{v}$. We write
$\ap\,\hat{f}\,\hat{e}\in\hat{f}\,\hat{u}\eqh\hat{f}\,\hat{v}$ where
$\hat{f}:\hat{A}\rah\hat{B}$ and $\hat{e}\in\hat{u}\eqh\hat{v}$, also
$\hat{\apd\,\hat{f}\,\hat{e} \in \TR_{\hat{P}}\,\hat{e}\,(\hat{f}\,\hat{u}) \eqh
  \hat{f}\,\hat{v}}$ where $\hat{\hat{f}\in(\hat{x}\in \hat{A})\rah \hat{B}}$
and $\hat{e} \in \hat{u}\eqh \hat{v}$. Borrowing notation from the homotopy type
theory book \cite{HoTTbook}, we write $\hat{p \sqcdot q}$ for transitivity and
$\hat{p^{-1}}$ for symmetry. We also make use of the groupoid law $\hat{\inv\,p
  : p^{-1}\sqcdot p = \refl}$.

@@

Note that this singleton contraction is not possible for a general non-strictly
positive function space. We rely on the domain being small: $(\underline{a})^\M$
yields an equation, but for general $\hat{\Gamma;\,}\Delta\vdash A$ types, $A^\M$ only yields
an unknown relation.

\subsection{The full homomorphism interpretation} We list below the $\blank^\M$
interpretations for the syntax. For brevity, we mostly leave $\hat{\gamma_0}$
and $\hat{\gamma_1}$ implicit, notating only $\hat{\gamma^M}$ where necessary.
We also omit some subscript parameters from $\hat{\J}$ applications.

\begingroup
\allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\M\hat{\gamma_0\,\gamma_1} && :\equiv \hat{\top} \\
  & (\Delta,x:A)^\M\,\hat{\gamma_0\,\gamma_1} && :\equiv \hat{(\gamma^M : \blc{\Delta^\M}\,\gamma_0\,\gamma_1)\times \blc{A^\M}\,\gamma^M} \\
  & x^\M\,\hat{\gamma^M} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^M} \\
  & \U^\M\,\hat{\gamma^M\,A_0\,A_1} && :\equiv \hat{A_0\ra A_1} \\
  & (\underline{a})^\M\,\hat{\gamma^M\,t_0\,\,t_1} && :\equiv \hat{
    \blc{a^\M}\,\gamma^M\,t_0 = t_1} \\
  & ((x:a)\ra B)^\M\,\hat{\gamma^M\,f_0\,f_1} && :\equiv \\
  & && \hat{\hspace{-8em}(x_0 : \blc{a^\A}\,\gamma_0)\ra \blc{B^\M}\,(\gamma_0,\,x_0)\,(\gamma_1,\,\blc{a^\M}\,\gamma^M\,x_0)\,(\gamma^M,\,\refl)\,
    (f_0\,x_0)\,(f_1\,(\blc{a^\M}\,\gamma^M\,x_0))}\\
  & (t\,u)^\M\,\hat{\gamma^M} && :\equiv  \hat{\J\,(\blc{t^\M}\,\gamma^M\,(\blc{u^\A}\,\gamma_0))\,(\blc{u^\M}\,\gamma^\M) }\\
  & ((\hat{x:A})\ra B)^\M\,\hat{\gamma^M\,f_0\,f_1} && :\equiv \hat{(x : A)\ra \blc{B^\M}\,\gamma^M\,(f_0\,x)\,(f_1\,x)} \\
  & (t\,\hat{u})^\M\,\hat{\gamma^M} && :\equiv t^\M\,\hat{\gamma^M\,u} \\
  & (t=_a u)^\M\,\hat{\gamma^M} && :\equiv
    \hat{\lambda\,e.\,(\blc{t^\M}\,\gamma^M\,^{-1})\sqcdot\ap\,(\blc{a^\M}\,\gamma^M)\,e\sqcdot\,\blc{u^\M}\,\gamma^M}\\
  & (\refl_t)^\M\,\hat{\gamma^M} && :\equiv \hat{\inv\,(\blc{t^\M}\,\gamma^M)}\\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\M\,\hat{\gamma^M} && :\equiv\\
  & && \hspace{-10em}\hat{\J\,\bigg(\J\,\Big(\J\,\big(\J\,(\lambda\,p_1\,p^M\,pr_1\,pr^M.\, pr^M)\,(\blc{t^\M}\,\gamma^M)}  \\
  & && \hspace{-9em}\hat{
      (\mathsf{uncurry}\,\blc{p^\A}\,\gamma_1)\,(\mathsf{uncurry}\,\blc{p^\M}\,\gamma^M)\,
      (\blc{pr^\A}\,\gamma_1)\,(\blc{pr^\M}\,\gamma^M)\big)\,
        (\blc{eq^\A}\,\gamma_0)\Big)(\blc{u^\M}\,\gamma^M)\bigg)(\blc{eq^\M}\,\gamma^M)   }\\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\M\,\hat{\gamma^M} && :\equiv\\
  & && \hspace{-9em}\hat{
    \J\,\big(\J\,(\lambda\,p_1\,p^M.\,\refl)\,(\blc{t^\M}\,\gamma^M)\,
      (\mathsf{uncurry}\,\blc{p^\A}\,\gamma_1)\,
      (\mathsf{uncurry}\,\blc{p^\M}\,\gamma^M)\big)
      (\blc{pr^\M}\,\gamma^M)
    }\\
  & ((\hat{x}\in \hat{A})\ra b)^\M\,\hat{\gamma^M} && :\equiv
       \hat{\lambda f\,x.\,\,\blc{b^\M}\,\gamma^M\,(f\,x)} \\
  & (t\,\hat{u})^\M\,\hat{\gamma^M} && :\equiv
       \hat{\ap\,(\lambda f.\,f\,u)\,(\blc{t^\M}\,\gamma^M)}
\end{alignat*}
\endgroup

@@

\[
\infer{\hat{\Gamma}\vdashh t^\D \in \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\hat{(}\hat{\gamma^D}\in\Delta^\D\,\hat{\gamma}\hat{)}\rah  A^\D\,\hat{\gamma}\,\hat{\gamma^D}\,\hat{(}t^\A\,\hat{\gamma}\hat{)}}{\hat{\Gamma}\semicol\Delta\vdash t : A}
\]
The implementation of $\blank^\D$ is given below. We leave
$\hat{\gamma}$-s mostly implicit, marking only $\hat{\gamma^D}$ witnesses.
\begingroup
%% \allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^D\,\hat{\gamma} && :\equiv \hat{\top} \\
  & (\Delta,\,x:A)^D\,(\hat{\gamma}\commah\,\hat{t}) && :\equiv \hat{(}\hat{\gamma^D}\in\Delta^D\,\hat{\gamma}\hat{)}\timesh A^D\,\hat{\gamma^D}\,\hat{t} \\
  & x^D\,\hat{\gamma^D} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^D} \\
  & \U^D\,\hat{\gamma^D}\,\hat{A} && :\equiv \hat{A} \rah  \Type_{\hat{i}} \\
  & (\underline{a})^D\,\hat{\gamma^D}\,\hat{t} && :\equiv a^\D\,\hat{\gamma^D}\,\hat{t} \\
  & ((x:a)\ra B)^D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\hat{(}\hat{x^D}\in a^\D\,\hat{\gamma^D}\,\hat{x}\hat{)} \rah  B^\D\,\hat{(}\hat{\gamma}\commah\hat{x}\hat{)}\,\hat{(}\hat{\gamma^D}\commah\hat{x^D}\hat{)}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,u)^D\,\hat{\gamma^D} && :\equiv \hat{(}t^\D\,\hat{\gamma^D}\hat{)}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\,\hat{(}u^\D\,\hat{\gamma^D}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  B^\D\,\hat{\gamma^D}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^D\,\hat{\gamma^D} && :\equiv t^\D\,\hat{\gamma^D}\,\hat{u} \\
  & (t=_a u)^D\,\hat{\gamma^D}\,\hat{e} && :\equiv \TR_{\hat{(}a^\D\,\hat{\gamma^D}\hat{)}}\,\hat{e}\,\hat{(}t^\D\,\hat{\gamma^D}\hat{)} \eqh u^\D\hat{\gamma^D} \\
  & (\refl_t)^D\,\hat{\gamma^D} && :\equiv \hat{\refl}_{\hat{(}t^\D\,\hat{\gamma^D}\hat{)}} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^D\,\hat{\gamma^D} && :\equiv \hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}pr^\D\,\hat{\gamma^D}\hat{)}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)}\hat{\big)}\,\hat{(}eq^\D\,\hat{\gamma^D}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^D\,\hat{\gamma^D} && :\equiv \hat{\refl} \\
  & ((\hat{x}\in \hat{A})\ra b)^D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  b^D\,\hat{\gamma^D}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^D\,\hat{\gamma^D} && :\equiv t^\D\,\hat{\gamma^D}\,\hat{u}
\end{alignat*}
\endgroup
The predicate for a context is given by iterating $\blank^\D$ for its
constituent types. For a variable, the corresponding witness is looked up from
$\hat{\gamma^D}$.

@@

Here, we mostly notate $\hat{\gamma^S}$ parameters and applications, making
$\hat{\gamma^D}$-s implicit in addition to $\hat{\gamma}$-s. The implementation
is the following.

@@

  & (\Delta,\,x:A)^\S\,(\hat{\gamma},\hat{t})\,(\hat{\gamma^D},\hat{t^D}) && :\equiv \hat{(}\hat{\gamma^S}\in\Delta^\S\hat{\gamma^2}\hat{)}\timesh A^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} \\
  & x^\S\,\hat{\gamma^S} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^S} \\
  & \U^\S\,\hat{\gamma^S}\,\hat{A}\,\hat{A^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  \hat{A^D}\,\hat{x} \\
  & (\underline{a})^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} && :\equiv a^\S\,\hat{\gamma^S}\,\hat{t} \eqh \hat{t^D} \\
  & ((x:a)\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\S\,\hat{(}\hat{\gamma,x}\hat{)}\,\hat{(}\hat{\gamma^D,}\,a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\,\hat{(}\hat{\gamma^S,}\,\hat{\refl}\hat{)} \\
  & && \hspace{8.4em} \hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{\big(}\hat{f^D}\,\hat{x}\,\hat{(}a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\hat{\big)} \\
  & (t\,u)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{(}t^\S\,\hat{\gamma^S}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\hat{)}\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah B^\S\,\hat{\gamma^S}\,\hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{(}\hat{f^D}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^S} && :\equiv t^\S\,\hat{\gamma^S}\,\hat{u} \\
  & (t=_a u)^\S\,\hat{\gamma^S}\,\hat{e} && :\equiv \TR\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\hat{\big(}\TR\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}\apd\,\hat{(}a^\S\,\hat{\gamma^S}\hat{)}\,\hat{e}\hat{)}\hat{\big)} \\
  & (\refl_t)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{\refl}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\S\,\hat{\gamma^S} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\bigg(}\hat{\J}\,\hat{\Big(}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\,\hat{pr^D}\,\hat{pr^S}.\,\hat{pr^S}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\,\\
  & && \hspace{-9em}\hat{(}\hat{\mathsf{uncurry}}\,p^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}pr^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}pr^\S\,\hat{\gamma^S}\hat{)}\hat{\big)}\,(eq^\A\,\hat{\gamma})\hat{\Big)}\,(u^\S\,\hat{\gamma^S})\hat{\bigg)}\,\hat{(}eq^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\S\,\hat{\gamma^S} && :\equiv \\
